### 2019-09-30 nodalida-talk/talk.tex
crucify gb4e
oops
corrections
additions

\documentclass[11pt]{beamer}
\usepackage{beamerthemesplit}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{gb4e}
\usepackage{tikz}
\usepackage{enumitem}
\usepackage{tabularx}
\usetikzlibrary{arrows,automata}
\usetheme{Boadilla}
% \usetheme{Warsaw}
\newcommand{\Sem}[1]{\sem{#1}}
\logo{\includegraphics{CLASP_logo.jpeg}} 
% \renewcommand{\theenumi}{\alph{enumi}}
\setitemize{label=\usebeamerfont*{itemize item}%
  \usebeamercolor[fg]{itemize item}
  \usebeamertemplate{itemize item}}
\newlist{lingex}{enumerate}{3}
\setlist[lingex,1]{parsep=0pt,itemsep=1pt,label=(\arabic*),resume=lingexcount}
\newcommand\onelingex[1]{\begin{lingex}\item #1 \end{lingex}}
\providecommand\ncases[1]{{\ensuremath{^{#1}}}}

% \newcommand{\ex}{}
% \newenvironment{exe}{}{}

@@

  \node[state]         (txt)                {text};
  \node[state]         (ast) [right of=txt] {AST};
  \node[state]         (pp)  [right of=ast] {Program};
  \node[state]         (res) [right of=pp] {result in $(0,1)$};

@@

  \path
  (Syntax) edge (GF)
  (Sem) edge (Hask)
  (Hyper) edge (Coq);
\end{tikzpicture}
}
  \begin{itemize}
  \item We use Lj\"unglof's FraCaS treebank and take these trees to
    their semantic counterparts
  \item GF constructions are translated to Haskell programs (types are
    respected). Dynamic semantics are handled at this level.
  \item The programs output Coq propositions and types. At the level
    of Coq, we add several kinds of linguistically relevant
    information (veridicality properties; adjectival semantics such as
    division into subsective, extentional, non-committal, etc.)
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{FraCoq: Dynamic Semantics}
  \begin{itemize}
  \item Monadic structure \cite{Shan:2002}
  \item Two kind of effects
    \begin{itemize}
    \item Manage a state of referents for anaphoric expressions (NPs,
      but also other linguistic categories)
    \item Non-determinism. (Linguistic expressions can have several
      interpretations.)
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{The FraCoq system: basic idea}
  \begin{itemize}
  \item Sentences are interpreted as propositions (with effect). 
  \item VP and CN are interpreted as predicates (also with effect).
  \item Adjectives: Functions from CN to CN (again: effect!)
    \begin{itemize}
    \item Different classes of adjectives are captured using coercions (subtyping). All special classes of adjectives are subtypes of A.
    \item Provision is made for intersective, subsective, privative and non-committal adjectives
    \end{itemize}
  \item (Generalized) Quantifiers. Similar to Montague semantics, but
    we also take into account quantities using an abstract function
    which can count the cardinalities of sets. (plus: effects, see below.)
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Anaphora}
  \begin{itemize}
  \item Wide-coverage: E-Type pronouns and Donkey anaphora, etc.
  \item New scope-extension mechanism
    \begin{lingex}
    \item
      \label{ex:anaphora-scope}
      \textbf{P1}	Every committee has a chairman.\\
      \textbf{P2}	He is appointed its members.\\
      \textbf{H} 	Every committee has a chairman appointed by members of the committee. \\ \textbf{Label:	YES} [FraCaS 122]
    \end{lingex}
    \begin{itemize}
    \item ``Every'' introduces a referent in the context.
    \item ``He'' find this referent.
    \item Normally, the referent would be out of scope, but we
      liberally extend scopes until we obtain a valid logical formula.
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Ellipsis}
  \begin{itemize}
  \item Similar to anaphora
  \item
    \begin{lingex}
    \item
      \label{ex:acda}
      \textbf{P1}	Bill spoke to everyone that John did [elliptic V2].\\
      \textbf{P2}	John spoke to Mary.\\
      \textbf{Q} 	Did Bill speak to Mary?\\
      \textbf{H} 	Bill spoke to Mary. \\ \textbf{Label:	Yes} [FraCas 173]
    \end{lingex}
  \item the verb ``spoke'' is introduced as a referent in the context when its encountered
  \item the argument of ``did'' is explicitly marked as an elliptic V2 by GF.
  \item we resolve it to ``speak'' by loading it from the environment.
  \end{itemize}
\end{frame}

\begin{frame}{Phrasal comparatives}
  \begin{itemize}
  \item Handled using the dynamic semantics
  \item Example:
    \begin{itemize}
    \item ``ITEL won more orders than APCOM lost.''
    \item becomes: ``APCOM lost some orders. ITEL won more
      orders [than some elliptic quantity].''
    \item Where is the elliptic quantity introduced?
    \item ``APCOM lost $\theta$ orders.  ITEL won at least $\theta$+1 orders.''
    \end{itemize}
  \end{itemize}
  (Simplified, see the paper for details.)
\end{frame}

@@

\begin{frame}
  \frametitle{Evaluation}
  \small
  \begin{tabularx}{9cm}{Xr@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r}
Section      & {\kern -2em} \#cases & FC2     & FC1 & MINE & Nut  & Langpro  \\ \hline
Quantifiers  & 75          & .96  & .96    & .77  & .53  & .93  \\
             &             & \ncases{74}     &        &      &      &     \ncases{44} \\
Plurals      & 33          & .82      & .76    & .67  & .52  & .73 \\
      &          &       &     &   &   & \ncases{24} \\
Anaphora     & 28          & .86      &   -    & -    & -    &  -       \\
Ellipsis     & 52          & .87      &   -    & -    & -    &  -       \\
Adjectives   & 22          & .95 & .95    & .68  & .32  & .73 \\
             &          &  \ncases{20} &     &   &   &  \ncases{12} \\
Comparatives {\kern -1em}& 31          & .87      & .56    & .48  & .45  &  -       \\
Temporal     & 75          &  -       &   -    &   -  &  -   &  -       \\
Verbs        & 8           & .75      &   -    & -    & -    &  -       \\
Attitudes    & 13          & .92      & .85    & .77  & .46  & .92  \\ 
    &          &       &     &   &   & \ncases {9}  \\ \hline
Total        & 337         & .89      & .83    & .69  & .50  & .85  \\
             &             & \ncases{259}    & \ncases{174}  & \ncases{174}& \ncases{174}& \ncases{89}
  \end{tabularx}

@@

@@

  \begin{itemize}
  \item	Improvement over earlier approaches. Still, there were a couple of difficulties, notably:
    \begin{itemize}
    \item Definite Plurals: Universal reading was captured. Cases of existential readings were not
      \begin{exe}
	\ex	
        \hskip 0.05in The inhabitants of Cambridge voted for a Labour MP.
        \hskip 0.05in very inhabitant of Cambridge voted for a Labour MP.
        \hskip 0.05in  	Did every inhabitant of Cambridge vote for a Labour MP? [UNK, FraCaS 094]			
      \end{exe}
    \item A syntactic conjunction of individuals is treated as a
      semantic group, or where precise counting of the members of a
      group is necessary.
    \end{itemize}\end{itemize}

>>>>

\documentclass[11pt]{beamer}
\usepackage{beamerthemesplit}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{tikz}
\usepackage{enumitem}
\usepackage{tabularx}
\usetikzlibrary{arrows,automata}
\usetheme{Boadilla}
% \usetheme{Warsaw}
\newcommand{\Sem}[1]{\sem{#1}}
%\logo{\includegraphics[height=0.7cm]{CLASP_logo.jpeg}} 
\logo{\includegraphics{clasp.ps}} 
\setitemize{label=\usebeamerfont*{itemize item}%
  \usebeamercolor[fg]{itemize item}
  \usebeamertemplate{itemize item}}
\newlist{lingex}{enumerate}{3}
\setlist[lingex,1]{parsep=0pt,itemsep=1pt,label=(\arabic*),resume=lingexcount}
\newcommand\onelingex[1]{\begin{lingex}\item #1 \end{lingex}}
\newcommand{\nologo}{\setbeamertemplate{logo}{}} % command to set the logo to nothing
\providecommand\ncases[1]{{\ensuremath{^{#1}}}}

\newcommand{\ex}{\item}
\newenvironment{exe}{\begin{lingex}}{\end{lingex}}

@@

  \node[state]         (txt)                {text};
  \node[state]         (ast) [right of=txt] {AST};
  \node[state]         (pp)  [right of=ast] {Program};
  \node[state]         (res) [right of=pp] {Result};

@@

  \path
  (Syntax) edge (GF)
  (Sem) edge (Hask)
  (Hyper) edge (Coq);
\end{tikzpicture}
}
  \begin{itemize}
  \item We use Lj\"unglof's FraCaS treebank and take these trees to
    their semantic counterparts
  \item GF constructions are translated to Haskell programs (types are
    respected). Dynamic semantics are handled at this level.
  \item The programs output Coq propositions and types. At the level
    of Coq, we add several kinds of linguistically relevant
    information (veridicality properties; adjectival semantics such as
    division into subsective, extensional, non-committal, etc.)
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{FraCoq 2.0: Dynamic Semantics}
  \begin{itemize}
  \item Monadic structure \cite{Shan:2002}
  \item Two kind of effects
    \begin{itemize}
    \item Manage a state of referents for anaphoric expressions (NPs,
      but also other linguistic categories)
    \item Non-determinism. (Linguistic expressions can have several
      interpretations.)
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{The FraCoq 2.0 system: basic idea}
  \begin{itemize}
  \item Sentences are interpreted as propositions (with effect). 
  \item VP and CN are interpreted as predicates (also with effect).
  \item Adjectives: Functions from CN to CN (again: effect!)
    \begin{itemize}
    \item Different classes of adjectives are captured using coercions (subtyping). All special classes of adjectives are subtypes of A.
    \item Provision is made for intersective, subsective, privative and non-committal adjectives
    \end{itemize}
  \item (Generalized) Quantifiers. Similar to Montague semantics, but
    we also take into account quantities using an abstract function
    which can count the cardinalities of sets. (plus: effects, see below.)
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Anaphora}
  \begin{itemize}
  \item Wide-coverage: E-Type pronouns and Donkey anaphora, etc.
  \item New scope-extension mechanism
    \begin{lingex}
    \item
      \label{ex:anaphora-scope}
      \textbf{P1}	Every committee has a chairman.\\
      \textbf{P2}	He is appointed by its members.\\
      \textbf{H} 	Every committee has a chairman appointed by members of the committee. \\ \textbf{Label:	YES} [FraCaS 122]
    \end{lingex}
    \begin{itemize}
    \item ``Every'' introduces a referent in the context.
    \item ``He'' finds this referent.
    \item Normally, the referent would be out of scope, but we
      liberally extend scopes until we obtain a valid logical formula.
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Ellipsis}
  \begin{itemize}
  \item Similar to anaphora
  \item
    \begin{lingex}
    \item
      \label{ex:acda}
      \textbf{P1}	Bill spoke to everyone that John did [elliptic V2].\\
      \textbf{P2}	John spoke to Mary.\\
      \textbf{H} 	Bill spoke to Mary. \\ \textbf{Label:	Yes} [FraCas 173]
    \end{lingex}
  \item The verb ``spoke'' is introduced as a referent in the context when its encountered
  \item The argument of ``did'' is explicitly marked as an elliptic V2 by GF.
  \item We resolve it to ``speak'' by loading it from the environment.
  \end{itemize}
\end{frame}

\begin{frame}{Phrasal comparatives}
  \begin{itemize}
  \item Handled using the dynamic semantics
  \item Example:
    \begin{itemize}
    \item ``ITEL won more orders than APCOM lost.''
    \item becomes: ``APCOM lost some orders. ITEL won more
      orders [than some elliptic quantity].''
    \item Where is the elliptic quantity introduced?
    \item ``APCOM lost $\theta$ orders.  ITEL won at least $\theta$+1 orders.''
    \end{itemize}
  \end{itemize}
  (Simplified, see the paper for details)
\end{frame}

@@

{\nologo
\begin{frame}
  \frametitle{Evaluation}
  \small
  \begin{tabularx}{9cm}{Xr@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r}
Section      & {\kern -2em} \#cases & FC2     & FC1 & MINE & Nut  & Langpro  \\ \hline
Quantifiers  & 75          & .96  & .96    & .77  & .53  & .93  \\
             &             & \ncases{74}     &        &      &      &     \ncases{44} \\
Plurals      & 33          & .82      & .76    & .67  & .52  & .73 \\
      &          &       &     &   &   & \ncases{24} \\
Anaphora     & 28          & .86      &   -    & -    & -    &  -       \\
Ellipsis     & 52          & .87      &   -    & -    & -    &  -       \\
Adjectives   & 22          & .95 & .95    & .68  & .32  & .73 \\
             &          &  \ncases{20} &     &   &   &  \ncases{12} \\
Comparatives {\kern -1em}& 31          & .87      & .56    & .48  & .45  &  -       \\
Temporal     & 75          &  -       &   -    &   -  &  -   &  -       \\
Verbs        & 8           & .75      &   -    & -    & -    &  -       \\
Attitudes    & 13          & .92      & .85    & .77  & .46  & .92  \\ 
    &          &       &     &   &   & \ncases {9}  \\ \hline
Total        & 337         & .89      & .83    & .69  & .50  & .85  \\
             &             & \ncases{259}    & \ncases{174}  & \ncases{174}& \ncases{174}& \ncases{89}
  \end{tabularx}

@@

\end{frame}		}

@@

  \begin{itemize}
  \item	Improvement over earlier approaches. Still, there were a couple of difficulties, notably:
    \begin{itemize}
    \item Definite Plurals: Universal reading was captured. Cases of existential readings were not

   \begin{lingex}
  	\item
  	\label{ex:acda}
  	\textbf{P1}	The inhabitants of Cambridge voted for a Labour MP.\\
  	\textbf{P2}	Every inhabitant of Cambridge voted for a Labour MP.\\
  	\textbf{H} Every inhabitant of Cambridge voted for a Labour MP.	\\ \textbf{Label:	UNK} [FraCas 094]
  \end{lingex}
    \item A syntactic conjunction of individuals is treated as a
      semantic group, or where precise counting of the members of a
      group is necessary.
    \end{itemize}\end{itemize}
